Computational methods in landscape ecology

### 2024-10-18 Sections/06-simulations.qmd
Making sure PR didnt add modelling

Individual-based models (or agent-based models) simulate discrete entities that are described by attributes and behavior.
Patterns emerge from bottom-up interactions of individuals with each other and their environment [@DeAngelis2014IndividualbasedModels].
In landscape ecology individual-based models are increasingly used to model social-ecological systems [@Carauta2018CanPreferential; @Dislich2018LanduseChange; @Gonzalez-Redin2019ExploringSustainable], but also disturbances [@Seidl2017ClimateChange; @Spies2017UsingAgentbased; @Kamimura2019AgentbasedModelling], or connectivity [@Day2020IndividualbasedModeling]. They can also incorporate a range of socio-economic, political and governance information to influence how agents (such as farmers) interact within a landscape [@murray2014combining].
In order to facilitate future model developments recent progress includes adaptable modelling frameworks [@Schumaker2018HexSimModeling] or reusable building blocks [@Berger2024ReusableBuilding].
Furthermore, hybrid approaches, coupling or integrating different model types with individual-based models, could benefit from rich development histories or facilitate corresponding strengths [@OSullivan2016StrategicDirections], e.g., linking an individual-based and ecosystem model [@Polhill2011AgentbasedModelling].

>>>>

Individual-based models (or agent-based models) simulate discrete entities that are described by attributes and behavior.
Patterns emerge from bottom-up interactions of individuals with each other and their environment [@DeAngelis2014IndividualbasedModels].
In landscape ecology individual-based models are increasingly used to model social-ecological systems [@Carauta2018CanPreferential; @Dislich2018LanduseChange; @Gonzalez-Redin2019ExploringSustainable], but also disturbances [@Seidl2017ClimateChange; @Spies2017UsingAgentbased; @Kamimura2019AgentbasedModelling], or connectivity [@Day2020IndividualbasedModeling]. 
They can also incorporate a range of socio-economic, political and governance information to influence how agents (such as farmers) interact within a landscape [@murray2014combining].
In order to facilitate future model developments recent progress includes adaptable modeling frameworks [@Schumaker2018HexSimModeling] or reusable building blocks [@Berger2024ReusableBuilding].
Furthermore, hybrid approaches, coupling or integrating different model types with individual-based models, could benefit from rich development histories or facilitate corresponding strengths [@OSullivan2016StrategicDirections], e.g., linking an individual-based and ecosystem model [@Polhill2011AgentbasedModelling].

### 2024-10-18 Sections/09-conclusion.qmd
Making sure PR didnt add modelling

Since its emergence in the 1980s [@With2019EssentialsLandscape], landscape ecology research has constantly evolved and is now highly dependent on computational methods.
Thus, future progress within the field will largely depend on integration of novel data sources and available software and computational tools [@Costanza2019DescribingAnalyzing].
Here, we present and summarize some recent developments of computational methods related to spatial patterns, connectivity, simulations and modelling, landscape genetics and spatial planning.
In this context, we highlight open-source software as a cornerstone of "Open Science" offering key advantages like shareability, reproducibility, and transparency, which provide great benefits to the research community [@Powers2019OpenScience; @Vogt2019PatternsSoftware].

>>>>

Since its emergence in the 1980s [@With2019EssentialsLandscape], landscape ecology research has constantly evolved and is now highly dependent on computational methods.
Thus, future progress within the field will largely depend on integration of novel data sources and available software and computational tools [@Costanza2019DescribingAnalyzing].
Here, we present and summarize some recent developments of computational methods related to spatial patterns, connectivity, simulations and modeling, landscape genetics and spatial planning.
In this context, we highlight open-source software as a cornerstone of "Open Science" offering key advantages like shareability, reproducibility, and transparency, which provide great benefits to the research community [@Powers2019OpenScience; @Vogt2019PatternsSoftware].
